---
title: "D-Lab Copilot Assisted Coding Workshop"
output: html_notebook
---

## Learning Objectives

Welcome to the D-Lab GitHub Copilot Workshop. After this workshop, you will be able to:

1.  Set-up and navigate Visual Studio Code
   
2.  Take advantage of the main functionalities of GitHub Copilot for coding
   
3.  Understand some of the strengths and weaknesses of AI coding assistants
  
Throughout this workshop series, we will use the following icons:

🔔 **Question**: A quick question to help you understand what's going on.

🥊 **Challenge**: Interactive exercise. We'll go through these in the workshop! Solutions to the challenges can be found in the solutions folder.

⚠️ **Warning**: Heads-up about tricky stuff or common mistakes.

💡 **Tip**: How to do something a bit more efficiently or effectively.

## 1. Getting Comfortable with Visual Studio Code

Visual Studio Code (VS Code) is a free code editor. When you start VS Code, you will be greeted by a Welcome screen that prompts you to start by opening a file or folder. Click `Open Folder` and open the GitHub-Copilot workshop folder that you downloaded from GitHub. 

Now that you have opened a folder, you will see an `Explorer` panel appear with all of the files within the folder you just opened. Click on the `workshop.Rmd` file to open it. You can open and close the `Explorer` panel by clicking on the icon that looks like two pieces of paper in the left-hand sidebar.

💡 **Tip**: When you are on your own computer in Visual Studio Code, you may notice the text sometimes goes off the screen if the line is very long. You can change your VS Code settings to make it so that the text in your markdown is automatically wrapped. To do this: 
1. Press `Ctrl + ,` or `Cmd + ,` to open the settings (You can also click on the cog icon in the bottom left corner of the window and then click `Settings`).
2. Search for word "wrap" in the search bar.
3. In the dropdown menu for `Editor: Word Wrap`, select on.
For more tips and tricks like this, check out the extensive and user-friendly Visual Studio Code documentation (https://code.visualstudio.com/docs) or ask you can even ask GitHub Copilot!

### Visual Studio Code Extensions
**Extensions** are add-ons that you can install to add additional features and tools to Visual Studio Code. You can find extensions by clicking on the square icon with four squares in the left-hand sidebar to open the `Extensions` panel. We will be using the following extensions in this workshop:
1. R - This extension provides support for the R programming language. You must have R already installed on your computer to use this extension.
2. GitHub Copilot - This extension is what gives us access to Copilot in VS Code

In the search bar, type R. Click on the R extension and then click the green `Install` button. Repeat for the GitHub Copilot extension. When you install the GitHub Copilot extension you will be prompted to login to your GitHub account. This is necessary to use Copilot.

### Running code in Visual Studio Code
Now that we have all the extensions we need, we can start running R code! To run a line of code, click on the line you want to run and press `Ctrl + Enter` or `Cmd + Enter`. To run multiple lines of code, highlight the lines you want to run and press `Ctrl + Enter`  or `Cmd + Enter`. If you are working in an R Markdown file, you can run a code chunk by clicking on the `Run Chunk` button that appears at the top of the chunk.

Run the R code below and see what happens.
```{r}
print("Yippee!")
```

When you run this code you will notice that a new window appears below: the `Terminal`. This is where the output of your code will appear. You can also run code in the terminal by typing directly into it and pressing `Enter`. If you have used RStudio before, the terminal is similar to the console in RStudio.

Now that we know how to run R, let's install and load the necessary packages for this workshop. First, let's install the `tidyverse` package. You can install it by running the code below. We will use this today for manipulating our data and for making plots.

```{r}
# If you haven't installed tidyverse before, uncomment this line 
# You may be prompted to pick a CRAN mirror: just choose the first one (0-Cloud [https])
#install.packages("tidyverse")

# If you have previously installed tidyverse, load it into the current R session
library(tidyverse)
```

We will also need the `httpgd` package to display plots interactively in VS Code. You can install it by running the code below.
```{r}
# if you haven't installed httpgd before, uncomment this line
#install.packages("httpgd")
```
💡 **Tip**: Unlike tidyverse or other packages you may actively use when coding, we do not need to load the `httpgd` package everytime (i.e., we don't need to run `library("httpgd")`). `httpgd` just provides the infrastructure for showing plots in VS Code.

Once we have installed the `httpgd` package, you need to set `httpgd` as the default plot viewer. You can do this by: 
1. Click the settings cog in the lower left hand corner or press `Ctrl + ,` or `Cmd + ,` to open the settings 
2. Search for `r.plot.useHttpgd` in the search bar
3. Click on the tick box to set it to true

Check that both tidyverse and httpgd are installed and working by running the code below:
```{r}
# Create a tiny data.frame
tiny_df <- data.frame(x = c(1, 2, 3), y = c(4, 5, 6))

# Plot the data
ggplot(tiny_df, aes(x = x, y = y)) +
  geom_point()
```

## 2. Introduction to GitHub Copilot

Now that we have set up Visual Studio Code, we can start using GitHub Copilot!

For our examples, we will be using data from [Gapminder](https://www.gapminder.org/), an educational non-proﬁt. It includes data for 142 countries, with values for life expectancy, GDP per capita, and population, every five years, from 1952 to 2007.

```{r}
# Read in data
gap <- read.csv(file = "data/gapminder.csv")
head(gap)
```

⚠️ **Warning**: By default, VS Code will set the working directory (i.e., where we "are" in your computer) to that of the folder you opened to start working in VS Code (in this case, the GitHub-Copilot folder). This is different than RStudio, where the working directory is set by default to the location of your file or R Project. You can check where your working directory is by running `getwd()`. Try it out below:
```{r}
getwd()
```

### GitHub Copilot Chat

GitHub Copilot chat allows you to directly ask questions to GitHub Copilot. To open the chat, cick on the icon that looks like two text bubbles on the left-hand sidebar. You can ask Copilot to generate code, explain code, or provide examples. You can also start a chat by pressing `Ctrl + Shift + I` or `Cmd + Shift + I`.  

💡 **Tip**: A common question we get is: what can Copilot "see" of my files and code? In general, Copilot can see code that you have highlighted or are working on in the open file. Copilot cannot see files that are not open or being used to interact with the chat (e.g., code outside the current file or session). Copilot also has limits on the amount of code it can process at once, so it may not be able to explain very long code chunks. However, it does learn from everything you show it, so the more you use it, the better it will get.

### 🥊 Challenge 1:
Open the chat, highlight the code in the cell below, and ask Copilot to explain it by typing a question into the chat (e.g.., "Explain this code, please"), no need to copy the code in, highlighting is enough. 

```{r}
gap <- read.csv(file = "data/gapminder.csv")
head(gap)
```

Copilot can also help us debug errors in our code. Below I have written some code to create histograms of the "lifeExp" and "gdpPercap" columns in the `gap` data frame. However, I have made three mistakes in the code. Can you find them? Highlight the code below and ask Copilot to help you debug it.
```{r}
ggplot(gap, aes(x = gdpPercap, y = lifeExp)) +
  geom_histogram(aes(fill = continet)) + 
  vline(xintercept = mean(gap$gdpPercap), color = "red") +
  labs(x = "GDP per Capita",  title = "Histogram of GDP per Capita")) +
  theme_minimal()
```

If you hover your cursor over the code chunk generated by Copilot you will see in the upper-right hand corner of the cell some buttons including (1) a `Copy` button that you can click to copy the code to your clipboard and (2) an `Insert at cursor` button that you can click to insert the code at the cursor in your open file. 

```{r}
# Add corrected code here
```

You may notice that in addition to fixing errors, Copilot also made some other improvements to the code. For example, in the response we got while making this exercise, it added a `binwidth` argument to the `geom_histogram` function to specify the width of the bins (see below). This `binwidth` argument is not necessary, but it can be helpful to make the plot more interpretable. Copilot even set the `binwidth` to 1000, which might be a reasonable choice for `gdpPercap`, but you should always check to make sure these changes make sense for your data.

```{r}
ggplot(gap, aes(x = gdpPercap)) +
  geom_histogram(aes(fill = continent), binwidth = 1000) +  # Corrected typo and added binwidth
  geom_vline(xintercept = mean(gap$gdpPercap), color = "red") +  # Corrected function name
  labs(x = "GDP per Capita", title = "Histogram of GDP per Capita") +  # Removed extra parenthesis
  theme_minimal() 
```

### In-line Chat
You can start an in-line chat with Copilot by pressing `Ctrl + I` or `Cmd + I`. This will allow you to ask Copilot for help with the code you are currently writing. When you activate in-line chat while highlighting code, Copilot will edit the code in the highlighted area.

### 🥊 Challenge 2: 
Highlight this ggplot code, press `Ctrl + I` or `Cmd + I`, and ask Copilot to change the background of the plot to your favorite color (e.g. "purple"), increase the text size of the x and y-axis titles to size 20, and make the title center aligned (you can highlight this prompt and copy it word for word into the inline chat). This is super helpful for when you can't remember what the exact syntax is for a specific ggplot2 element.

Click "Accept" to accept the changes Copilot suggests. You can also click "Discard" to discard the changes, "Rerun Request" to ask Copilot to generate new code, or "Toggle changes" to see the changes Copilot made (the "Toggle Changes" option may appear in the dropdown menu which you can access by clicking the down arrow next to the "Discard" button).

Accept the changes and run the chunk to see your new plot. 

```{r}
ggplot(gap, aes(x = gdpPercap)) +
  geom_histogram(binwidth = 1000, fill = "blue", color = "black") +
  labs(x = "GDP per Capita", title = "Histogram of GDP per Capita")
```

⚠️ **Warning**: Sometimes you may notice that Copilot doesn't do everything you told it to. This is because Copilot is still learning and may not always understand what you are asking for. You can always ask it again or make the changes yourself. For example, when you asked to change the background, it may have changed the background of the plot, when what you really wanted to do is to change the background of the panel, or vise versa; in that case, you would have to update your request to specify that you want to change the panel background.

### In-line suggestions
Another handy feature of Copilot is in-line suggestions as you type. These suggestions appear in a lighter gray text after your cursor and you can accept them by pressing `Tab`. These suggestions can be very helpful for completing code quickly. Suggestions are automatically triggered by the code your write, based on context from the code you have already written. You may notice as you use Copilot more often that the suggestions even learn your "style" of coding.

### 🥊 Challenge 3: 
Start typing `unique_countries <- gap %>%` and see what suggestions Copilot gives you. You may need to wait a second for the suggestion to appear or continue typing. Click `Tab` to accept the suggestion. This should complete the code to create a vector of unique countries in the `gap` data frame.
```{r}
# Start typing here
```

⚠️ **Warning**: Copilot sometimes provides suggestions that are not what you want. You can always ignore the suggestion and continue typing. Copilot can change over time and give different suggestions based on even minor differences in context clues. You can test this out by deleting the line of code you just wrote and typing it again to see if you get the same suggestion.

### 🥊 Challenge 4:
You can also trigger suggestions by providing comments in your code. These commments allow you to give more detailed information and context for the in-line suggestions. Start typing `# Calculate average gdpPercap and lifeExp grouped by continent for the year 2007` and see what suggestions Copilot gives you. Click `Tab` to accept the suggestion.

💡 **Tip**: You may have to start writing `gap %>%` to trigger the suggestions.
```{r}
# Start typing here
```

Now you have seen how we can use both the chat, in-line chat, and in-line suggestions to interact with Copilot. So far we have used some simple examples, but Copilot can also help with more complex code.

### 🥊 Challenge 5: 
Here, we have code below that is a bit untidy (missing indents, no comments, long single lines of code). Highlight the code below and in the chat (you can use the in-line or the regular chat) ask Copilot to reformat this code and add comments. 
```{r}
gap_mean <- gap%>%filter(continent=="Asia")%>%group_by(year)%>%summarise(mean = mean(lifeExp))
ggplot(gap_mean, aes(x = year, y = mean)) +
geom_line() +
theme_classic() +
labs(x = "Year", y = "Life Expectancy", title = "Average Life Expectancy in Asia by Year")

gap_mean <- gap%>%filter(continent=="Europe")%>%group_by(year)%>%summarise(mean = mean(lifeExp))
ggplot(gap_mean, aes(x = year, y = mean)) +
geom_line() +
theme_classic() +
labs(x = "Year", y = "Life Expectancy", title = "Average Life Expectancy in Europe by Year")

gap_mean <- gap%>%filter(continent=="Oceania")%>%group_by(year)%>%summarise(mean = mean(lifeExp))
ggplot(gap_mean, aes(x = year, y = mean)) +
geom_line() +
theme_classic() +
labs(x = "Year", y = "Life Expectancy")
```

You may notice that you have to do some extra work to complete the clean-up; for example, you may want to remove unnecessary or incorrect comments or add titles that are missing.

```{r}
# Paste new cleaned code here
```

We can take this clean-up a step further and ask it to make a function from our repetitive code. Highlight the code above and ask Copilot to reduce repetitiveness in the code by turning it into a function that takes in data and a continent name (for some of you, it may have done this automatically in your first try based on how you asked for the clean-up)

```{r}
# Paste the new function code here
```

A common pitfall when working with Copilot is using it to generate code without understanding what the code is doing. **Just because code runs without an error does not mean it is correct.** Remember, you are the pilot and Copilot is...the copilot; it is crucial to use your knowledge and expertise to guide your coding decisions. To demonstrate this pitfall, let's ask Copilot to do something very broad: figure out what drives life expectancy.

```{r}
# Using the chat, ask Copilot to determine the driver of life expectancy in our data
```

You will probably get some code that builds a linear model using the lm() function. For example:

```{r}
mod <- lm(lifeExp ~ ., data = gap)
summary(mod)
```

What next steps would you take to assess the validity of this model?

Next, let's check the residuals of the model to see if they are normally distributed:
```{r}
# Q-Q plot for residuals
qqnorm(residuals(mod))
qqline(residuals(mod), col = "red")
```

The residuals do not appear to be normally distributed (i.e., the points deviate from the Q-Q line). Let's do some additional data visualization to figure out what is going on:
```{r}
ggplot(gap, aes(x = pop, y = lifeExp)) +
  geom_point() +
  geom_smooth(method = "lm") + 
  theme(legend.position = "none") 

ggplot(gap, aes(x = year, y = lifeExp)) +
  geom_smooth(method = "lm") + 
  geom_point() +
  theme(legend.position = "none")

ggplot(gap, aes(x = gdpPercap, y = lifeExp)) +
  geom_smooth(method = "lm") + 
  geom_point() +
  theme(legend.position = "none")
```

We notice that the relationship between some of the variables and life expectancy are non-linear. We are going to focus on population size, because there is a particularly strong non-linear pattern in the data and it appears as if we may need an interaction term. Let's plot the points colored by continent and country to see if there is a pattern:

```{r}
ggplot(gap, aes(x = pop, y = lifeExp, color = continent)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme(legend.position = "none")

ggplot(gap, aes(x = pop, y = lifeExp, color = country)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme(legend.position = "none")
```

We can see that the relationship varies by country and fitting one line to all of the countries is probably not a good idea. If this was a real analysis, there are  additional steps we would take to address this issue, but the point of this exercise is that everything we just did to validate our model was not suggested by Copilot when we asked it to identify the drivers of life expectancy. If we had just gone along with what Copilot gave us, we may not have discovered the problem with population size. **Not all incorrect code will result in a error or warning, the most common and insidious code problems are silent**, which is why it is essential to understand what your code is doing. We had to use our knowledge as data scientists to guide our analysis. Copilot is still helpful for creating code to run our models, generate plots, etc., but we need to be the ones in the drivers seat. You are the expert!

## Key Points

Summary of what was learned in today's workshop:

-   Visual Studio Code is a code editor that has extensions to allow you to code in various langauges (i.e., R) and to use GitHub Copilot.

-   To chat with Copilot, you can use the chat icon in the left-hand sidebar or use the in-line chat feature (`Ctrl + I` or `Cmd + I`).
  
-   Copilot can provide suggestions as you type, based on the context of your code. Press `Tab` to accept a suggestion.

-   Copilot can help you clean-up your code and change repetitive code into functions.

-   Copilot is still learning and may not always provide the exact code you want. You may need to make changes to the code it provides. The more you use Copilot, the better it will get.

💡 **Tip**: If you have been going through this workshop in our GitHub Codespace and you want to download your edited file, right-click on the file you want to download (i.e. lessons/workshop.Rmd) in the `Explorer` panel on the left-hand side and click `Download...`  (Note: you may have to click `Allow` on a security pop-up). If you are comfortable with Git, you can also fork this repository and commit your changes to your fork (instructions [here](https://docs.github.com/en/codespaces/getting-started/understanding-the-codespace-lifecycle#saving-changes-in-a-codespace)).
