---
title: "D-Lab GitHub Copilot Workshop"
output: html_notebook
---
```{r}
# Load the tidyverse
library(tidyverse)

# Read in data
gap <- read.csv(file = "data/gapminder.csv")
head(gap)
```

### 🥊 Challenge 1:
Open the chat, highlight the code in the cell below, and ask Copilot to explain it. Then ask Copilot to give you code to generate a summary of the data in the `gap` data frame by plotting histograms of "lifeExp" and "gdpPercap".

This is the chat response we got. note that your chat response may be different:
"This code is written in R and it's doing two things:

gap <- read.csv(file = "data/gapminder.csv"): This line is reading a CSV file named "gapminder.csv" located in a directory named "data". The data from the file is being loaded into a variable named gap.

head(gap): This line is displaying the first six rows of the data frame gap. The head() function in R is used to show the first part of objects like vectors, tables, data frames, etc. By default, it shows the first 6 rows."

```{r}
gap <- read.csv(file = "data/gapminder.csv")
head(gap)

# Load the necessary library
library(ggplot2)

# Plot histogram for lifeExp
ggplot(gap, aes(x=lifeExp)) +
  geom_histogram(binwidth=5, fill="#69b3a2", color="#e9ecef", alpha=0.9) +
  theme_minimal() +
  labs(title="Histogram of Life Expectancy", x="Life Expectancy", y="Count")

# Plot histogram for gdpPercap
ggplot(gap, aes(x=gdpPercap)) +
  geom_histogram(binwidth=1000, fill="#69b3a2", color="#e9ecef", alpha=0.9) +
  theme_minimal() +
  labs(title="Histogram of GDP per Capita", x="GDP per Capita", y="Count")
```


### 🥊 Challenge 2: 
Highlight this ggplot code, press `Ctrl + I`, and ask Copilot to change the background of the plot to your favorite color (e.g. "purple"), increase the text size of the x and y-axis titles to size 20, and make the title center aligned. This is super helpful for when you can't remember what the exact syntax is for a specific ggplot2 element.
```{r}
ggplot(gap, aes(x = gdpPercap)) +
  geom_histogram(binwidth = 1000, fill = "blue", color = "black") +
  labs(x = "GDP per Capita", title = "Histogram of GDP per Capita") +
  theme(plot.background = element_rect(fill = "aquamarine"),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        plot.title = element_text(hjust = 0.5))
```

### 🥊 Challenge 3: 
You can also use in-line chat in a blank line to generate new code. Start in-line chat in the empty chunk below and ask Copilot to generate code to calculate the mean and variance of the "pop" column in the `gap` data grouped by the "country" column.
```{r}
mean_var <- gap %>%
  group_by(country) %>%
  summarise(mean_pop = mean(pop), var_pop = var(pop))

mean_var
```

### In-line suggestions
You may have noticed throughout this workshop that Copilot provides suggestions as you type in a lighter gray text. These suggestions can be helpful for completing code quickly. You can accept a suggestion by pressing `Tab`. 

Suggestions are automatically triggered by the code your write, based on context from the code you have already written. 

### 🥊 Challenge 4: 
Start typing `unique_countries <- gap` and see what suggestions Copilot gives you. Accept the suggestion that completes the code to create a vector of unique countries in the `gap` data frame.
```{r}
unique_countries <- 
  gap %>% 
  select(country) %>% 
  unique()
```

### 🥊 Challenge 5:
You can also trigger suggestions by providing comments in your code. These commments allow you to give more detailed information and context for the in-line suggestions. Start typing `# Calculate average gdpPercap and lifeExp grouped by continent for the year 2007` and see what suggestions Copilot gives you. 

💡 **Tip**: You may have to start writing `gap %>%` to trigger the suggestions.
```{r}
# Calculate average gdpPercap and lifeExp grouped by continent for the year 2007
gap_2007 <- gap %>%
  filter(year == 2007) %>%
  group_by(continent) %>%
  summarise(mean_gdpPercap = mean(gdpPercap), mean_lifeExp = mean(lifeExp))
```

### 🥊 Challenge 6: 

Copilot can be really useful for cleaning up your code. For example, we have code below that is a bit untidy (missing indents, no comments, long single lines of code). Highlight the code below and in the chat ask Copilot to format this code and add comments.
```{r}
gap_mean <- gap%>%filter(continent=="Asia")%>%group_by(year)%>%summarise(mean = mean(lifeExp))
ggplot(gap_mean, aes(x = year, y = mean)) +
geom_line() +
theme_classic() +
labs(x = "Year", y = "Life Expectancy", title = "Average Life Expectancy in Asia by Year")

gap_mean <- gap%>%filter(continent=="Europe")%>%group_by(year)%>%summarise(mean = mean(lifeExp))
ggplot(gap_mean, aes(x = year, y = mean)) +
geom_line() +
theme_classic() +
labs(x = "Year", y = "Life Expectancy", title = "Average Life Expectancy in Europe by Year")

gap_mean <- gap%>%filter(continent=="Oceania")%>%group_by(year)%>%summarise(mean = mean(lifeExp))
ggplot(gap_mean, aes(x = year, y = mean)) +
geom_line() +
theme_classic() +
labs(x = "Year", y = "Life Expectancy")
```

You may notice that you have to do some extra work to complete the clean-up; for example, you may want to remove unnecessary or incorrect comments or add titles that are missing.

```{r}
# Calculate the mean life expectancy for Asia by year
gap_mean <- gap %>%
  filter(continent == "Asia") %>%
  group_by(year) %>%
  summarise(mean = mean(lifeExp))

# Plot the mean life expectancy for Asia by year
ggplot(gap_mean, aes(x = year, y = mean)) +
  geom_line() +
  theme_classic() +
  labs(x = "Year", y = "Life Expectancy", title = "Average Life Expectancy in Asia by Year")

# Calculate the mean life expectancy for Europe by year
gap_mean <- gap %>%
  filter(continent == "Europe") %>%
  group_by(year) %>%
  summarise(mean = mean(lifeExp))

# Plot the mean life expectancy for Europe by year
ggplot(gap_mean, aes(x = year, y = mean)) +
  geom_line() +
  theme_classic() +
  labs(x = "Year", y = "Life Expectancy", title = "Average Life Expectancy in Europe by Year")

# Calculate the mean life expectancy for Oceania by year
gap_mean <- gap %>%
  filter(continent == "Oceania") %>%
  group_by(year) %>%
  summarise(mean = mean(lifeExp))

# Plot the mean life expectancy for Oceania by year
ggplot(gap_mean, aes(x = year, y = mean)) +
  geom_line() +
  theme_classic() +
  labs(x = "Year", y = "Life Expectancy", title = "Average Life Expectancy in Oceania by Year")
```

We can take this clean-up a step further and ask it to make a function from our repetitive code. Highlight the code above and ask Copilot to turn it into a function. 

```{r}
# Function to calculate and plot mean life expectancy by year for a given continent
plot_life_expectancy <- function(continent) {
  # Calculate the mean life expectancy for the given continent by year
  gap_mean <- gap %>%
    filter(continent == continent) %>%
    group_by(year) %>%
    summarise(mean = mean(lifeExp))

  # Plot the mean life expectancy for the given continent by year
  ggplot(gap_mean, aes(x = year, y = mean)) +
    geom_line() +
    theme_classic() +
    labs(x = "Year", y = "Life Expectancy", title = paste("Average Life Expectancy in", continent, "by Year"))
}

# Call the function for Asia
plot_life_expectancy("Asia")

# Call the function for Europe
plot_life_expectancy("Europe")

# Call the function for Oceania
plot_life_expectancy("Oceania")
```

We can also prompt Copilot to generate a function from scratch. You can do this by either using the chat, prompting with a comment (i.e., `# Create a function that...`), or just starting to write a function (i.e., `function_name <- function(...)`; in this case, Copilot will use the name of the function to guess at what you are trying to make). Try to create your own custom function below by using this last approach. For example, you could prompt Copilot to return the colors of the rainbow by writing `rainbow_colors <- function()`.

💡 **Tip**: You may have to use some comments (e.g., `# Return a vector of colors`) to give more context to Copilot for crafting your function.
```{r}
# Function returned by Copilot:
rainbow_colors <- function() {
  # Return a vector of colors
  colors <- c("red", "orange", "yellow", "green", "blue", "indigo", "violet")
  return(colors)
}

# Note that "indigo" is not a recognized color in R, but it is included here for the sake of the rainbow sequence. Below, we have removed indigo and plotted the colors.

rainbow_colors <- function() {
  # Return a vector of colors
  colors <- c("red", "orange", "yellow", "green", "blue", "violet")
  return(colors)
}

# ggplot of the rainbow colors 
ggplot(data.frame(x = 1:6, y = 1:6), aes(x = x, y = y, color = rainbow_colors())) +
  geom_point(size = 5) +
  scale_color_identity()
```